Volatility and the Buyback Anomaly: Interactive Tool

### 2017-02-17 manuscript.Rnw
clean up

if (generate_all_data == 1){
  source("create_bbissuers_data.R")
}
if (generate_rnw_data == 1){
  source("bb_issuers_new.R")
} else {
  load("bb_issuers_new.Rdata")
}

@@

\doublespacing
% Use endnotes instead of footnotes - redefine \footnote command
%\renewcommand{\footnote}{\endnote}  % Endnotes instead of footnotes

@@

%Why does the \cite{StambaughYuan2016} four factor model does a poor job explaining the buyback anomaly? The reason, we believe, is that the buyback anomaly is inconsistent with some of the major 11 anomalies used to construct the \cite{Stambaugh2015} mispricing index. For example, long-term excess returns after buyback announcements are negatively related to the returns in the 6 months prior to the buyback. As buyback stocks have negative momentum before the announcement, the momentum anomaly \citep{JegadeeshTitman1993} predicts lower returns not higher returns. Similarly, \cite{PeyerVermaelen2009} find that the most undervalued buyback stocks miss earnings forecast, which is the opposite of the \cite{Mendenhall2004} earnings drift anomaly which predicts negative excess returns after missing earnings forecasts.

@@

%The prediction of the costly arbitrage hypothesis is that there is a positive relation between volatility and excess returns after buybacks and a negative relation between volatility and excess returns after equity issues. While the first hypothesis is confirmed by the data, the second one is not. The results are more consistent with the options hypothesis of \cite{IkenberryVermaelen1996} who point out that open market repurchase authorizations are options to repurchase stock, not firm commitments. As a result, they predict that announcement returns should be positively related to volatility.

%Moreover, because limits to arbitrage are higher among overvalued stocks (e.g. most investors are reluctant or unable to short stocks) the negative volatility effect among overvalued stocks should be stronger than the positive volatility effect among undervalued stocks. Indeed, \cite{Bravetal2010} find that the limits to arbitrage can explain some overvaluation anomalies but not undervaluation anomalies. So whether the costly arbitrage hypothesis can explain the buyback anomaly and/or the new equity issue anomaly is an empirical question.  
%The prediction of the costly arbitrage hypothesis is that there is a positive relation between volatility and excess returns after buybacks and a negative relation between volatility and excess returns after equity issues. While the first hypothesis is confirmed by the data, the second one is not. The results are more consistent with the options hypothesis of \cite{IkenberryVermaelen1996} who point out that open market repurchase authorizations are options to repurchase stock, not firm commitments. As a result, they predict that announcement returns should be positively related to volatility. Note that as equity issues are firm commitments, no such volatility effect is predicted. We also argue that the market timing hypothesis implicitly assumes that managers have superior knowledge about company-specific information, not the overall market. Hence the prediction of the information advantage hypothesis is that buyback announcements of firms with high standardized idiosyncratic volatility, measured by (1-$R^2$), will also generate higher long term excess returns. Indeed, it measures to what extent the volatility of stock returns is explained by company-specific (non-factor related) information [see also \citep{LiRajgopaletal2014}].

@@

%The fact that the buyback anomaly survives could still mean that the ''excess'' returns are compensating for some risks that the Fama-French 5-factor model does not account for - and is not related to any information advantage of company insiders. The third contribution of this paper is to provide further support for the hypothesis that excess returns after buybacks can be explained by information asymmetry rather than a risk-based story. One way to show that these excess returns reflect adjustment to undervaluation is to show that they are positively correlated with observable indicators of potential undervaluation and/or information advantage of the management. \cite{IkenberryLakonishokVermaelen1995} assume that value stocks are more likely to be undervalued than growth stocks. The fact that they find a positive relation between long-term excess returns and book-to-market supports the undervaluation hypothesis.  The undervaluation index of \cite{PeyerVermaelen2009} assumes that buybacks announced by beaten up small cap value firms are more likely to be driven by undervaluation than other firms. consider three new indicators (firm size, prior return and stated motivation in the press release) and combine them with the book-to-market indicator in an undervaluation index (U-index). Adding these additional indicators improves the predictability of excess returns, which is consistent with the undervaluation hypothesis. In this paper we test whether two new indicators, standardized idiosyncratic volatility (measured by 1-$R^2$) and volatility can improve our understanding of excess returns by combining them with the U-index of \cite{PeyerVermaelen2009}.

%It remains a fact though that the buyback anomaly is to some extent a small firm anomaly, as also found by \cite{PeyerVermaelen2009}: indeed, we find that value-weighting all the events [as suggested by \cite{MitchellStafford2000}] makes the (Calendar Time event study method) alphas disappear. However, if we exclude the firms in the largest size quartile, the anomaly persists (also for the Calendar Time event study method).   This makes sense as small firms are more opaque and less followed by analysts. They also tend to be held more by insiders, so that in these case share buybacks are more similar to insider buying with company funds. If anything, to increase the power of the test to detect mispricing and the possibility that managers (at least in some firms) can time the market based on their information advantage relative to investors, any weighting should be based on the inverse of size. The fact that excess returns disappear when we value weight provides further support for the hypothesis that excess returns are driven by information asymmetry for some firms. The fact that this is a small cap anomaly may partly explain why this anomaly persists after 30 years and has attracted very little attention in the asset management industry.\footnote{For example, a Google search for ``buyback funds'' gives very few results: Powershare Buyback Achievers fund, KBC Buyback America, S\&P 500 Buyback ETF, Catalyst/Equity Compass Buyback Strategy fund, and PV Buyback USA. The first 3 funds focus on large caps after buyback completions although the academic research shows abnormal returns are more significant in small, under-priced, value stocks and the relevant event is not the completion of the buyback but the buyback authorization. We are also not aware of event-driven hedge funds that buy repurchasing firms and short equity issuers; typical event-driven strategies are for example based on M\&A arbitrage, capital structure arbitrage or on investing in distressed securities.}  Indeed because management fees are proportional to fund size one expects relatively less interest in anomalies concentrated in small caps or microcaps.

@@

%Finally, we show that adding volatility and standardized idiosyncratic volatility to the factors proposed by previous research as undervaluation proxies and combining them in an Enhanced Undervaluation index improves our forecasting ability of excess returns after buyback authorization announcements. As a theoretical argument can be made that both measures are proxies for the competitive information advantage of managers-insiders, this strengthens the support for the undervaluation and market timing hypothesis. However, it should be noted that the anomaly is more pronounced in small firms, which may explain why this anomaly has attracted little attention in the asset management industry.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%% OLD INTRO

%We study whether previously reported positive (negative) excess returns following share buyback (equity issue) announcements\footnote{For evidence on long-term excess returns after buybacks see e.g., \cite{IkenberryLakonishokVermaelen1995,PeyerVermaelen2009,ManconiPeyerVermaelen2015}. For evidence  on under-performance after equity issues see e.g., \cite{LoughranRitter1995,SpiessAffleckGraves1995,Eckboetal2000,DittmarThakor2007,Bravetal2000}.} are compensation for risk using recently proposed risk models [specifically from \cite{FamaFrench2015} or \cite{Houetal2015}], or evidence for possible market timing by insiders of undervalued (overvalued) firms who may take advantage of information asymmetries. We also explore the impact of firm specific or total volatility on excess returns, as theoretical arguments can be made that high volatility makes arbitrage for outsiders costlier. Specifically, the first purpose of this paper is to test whether there is a link between two well-known anomalies: the net issue anomaly, i.e. the fact that share buybacks (equity issues) are followed by long run positive (negative) excess returns, and the volatility anomaly, i.e. the fact that (idiosyncratic) volatility is negatively related to expected returns \citep{Ang2006}. This link is plausible considering the fact that \cite{Stambaugh2015} find that the relation between idiosyncratic volatility, measured as residual variance, and future returns reverses and becomes {\it positive} for undervalued stocks.\footnote{They  define undervaluation on the basis of a combination of 11 return anomalies reported in the literature, including net equity issuance.} They argue that their result is consistent with the {\it costly  arbitrage hypothesis}: idiosyncratic volatility represents risk that deters arbitrage and the resulting reduction in mispricing, hence among underpriced stocks, the stocks with the highest idiosyncratic volatility should be the most underpriced. As the positive (negative) long run returns after buybacks (equity issues) are generally interpreted as evidence of undervaluation (overvaluation) that firm insiders may take advantage of, we would predict also a positive (negative) relation between volatility and future returns after firms announce a buyback (equity issue): volatility may also make it easier for insiders to take advantage of undervaluation (overvaluation) based on an information advantage on the one hand and the challenge for outsiders to correct mispricing due to higher arbitrage costs on the other hand.

%Moreover, \cite{Stambaugh2015} argue that, because limits to arbitrage are higher among overvalued stocks (e.g. most investors are reluctant or unable to short stocks) the negative volatility effect among overvalued stocks should be stronger than the positive volatility effect among undervalued stocks.  Indeed, \cite{Bravetal2010} find that the limits to arbitrage can explain some overvaluation anomalies but not undervaluation anomalies such as the fact e.g. that positive earnings surprises are followed by positive excess returns.  So whether the costly arbitrage hypothesis is also related to the buyback undervaluation anomaly is ultimately an empirical question.

%The second purpose of this paper is to test whether the buyback and equity issue anomalies still exist and are not simply a proxy for risk, or have disappeared in recent years as has happened with other ones~\citep{McLean2016}. Unlike past work, we do so by both using the more recent \cite{FamaFrench2015} five-factor  model, for buyback and equity announcements during the period \Sexpr{firstyear}-\Sexpr{lastyear}, and using event announcement dates and separating equity issues and buybacks.\footnote{Using throughout this paper the q-factor model of \cite{Houetal2015} instead of the five-factor model of \cite{FamaFrench2015} leads to the same conclusions. Results available upon request. } Excess returns in previous research are calculated using different benchmarks such as firms with similar size and book-to-market ratio \citep{IkenberryLakonishokVermaelen1995} or the \cite{FamaFrench1993} three-factor model or the \cite{Carhart1997} 4-factor model \citep{PeyerVermaelen2009}. However, \cite{FamaFrench2016}  argue that many anomalies are weakened or do not survive after using the more recent \cite{FamaFrench2015} five-factor asset pricing model as a model of expected returns. This model incorporates new evidence that profitability and investment patterns, besides market to book and size, explain stock returns \citep{NovyMarx2013}. They find that the volatility anomaly survives for small firms, but the net issue anomaly (i.e. equity issues minus buybacks) does not. If buybacks (equity issues) are done by firms with high (low) profitability and few (many) investment opportunities, then these factors may well explain the excess returns reported in previous research.

%Moreover, \cite{FamaFrench2016} do not exactly replicate the papers that first reported the anomalies. Indeed, unlike the authors that ''discovered'' the anomalies they do not examine repurchases and equity issues separately: they calculate returns after net equity issues (funds spent on buybacks minus funds spent on equity issues). The part of their sample where net equity issues are positive is defined as the ''buyback sample''. But this sample still contains some equity issuers, which may introduce a downward bias in the excess returns calculations.  Moreover, they assume investors buy after the completion of the buyback and the equity issue, not around the announcement date of the buyback authorization as is done in previous research [e.g., \cite{PeyerVermaelen2009}]. For buybacks this may be an issue as repurchases may be completed several years after the buyback authorization, or not completed at all \citep{StephensWeisbach1998}. This slow execution is partially driven by SEC regulations designed to minimize the impact on trading volume as well as tactical considerations. This may produce another downward bias in post-repurchase excess returns. For example, assume the stock price today is \$10 and the management believes its stock is undervalued and announces a typical buyback program for 7$\%$ of its outstanding shares. Suppose that one year later 2$\%$ of the shares are repurchased and the market corrects the undervaluation so that the company halts the program.  At the end of the first year the firm is considered as a net repurchaser in the \cite{FamaFrench2016} sample, and no long-term excess return will be observed afterwards even though there were excess returns during the first post-announcement year.

%Pooling buybacks and equity issues in a ''net issues'' measure assumes that the decision to issue equity is simply the mirror of buying back shares. This ignores two other major differences between repurchases and equity issues (besides the fact that a buyback program is an option, not a firm commitment as an equity issue).  First, in an open market repurchase the seller is not aware she is selling to the corporation, while in an equity issue the investor knows that the company is the issuer. Obviously this makes it easier to buy back undervalued shares than to issue overvalued shares. Second, issuing overvalued shares to new investors may hurt relations with these investors. One of the arguments for IPO underpricing is that it creates a positive experience for investors making it easier to convince them to buy new shares in a follow up secondary offering \citep{Ibbotson1975}. Issuing overvalued stock creates a negative experience for new investors. On the other hand, repurchasing undervalued shares from investors who have decided to sell their shares anyway should not create an investor relations problem. So in order to test the buyback and equity issue anomalies one has to examine them separately, using announcement dates as we do in this paper.

%We confirm the \cite{FamaFrench2016} conclusion that the five-factor model makes the equity issue anomaly disappear, but the buyback anomaly remains statistically and economically significant. The buyback anomaly is also persistent over time and does not seem to become less significant in recent years, which is inconsistent with the hypothesis that the growth of institutional investors and the reduction in trading costs may have made markets more efficient as argued by \cite{FuHuang2016}.  As we find that only the buyback anomaly survives, this paper focuses on buybacks, not equity issues.

%The fact that the buyback anomaly survives could still mean that the ''excess'' returns are compensating for some risks that the Fama-French 5-factor model does not account for - and is not related to any information advantage of company insiders. The third contribution of this paper is to provide further support for the hypothesis that excess returns after buybacks can be explained by information asymmetry rather than a risk-based story. One way to show that these excess returns reflect adjustment to undervaluation is to show that they are positively correlated with observable indicators of potential undervaluation and/or information advantage of the management.  \cite{IkenberryLakonishokVermaelen1995} assume that value stocks are more likely to be undervalued than growth stocks. The fact that they find a positive relation between long-term excess returns and book-to-market supports the undervaluation hypothesis. \cite{PeyerVermaelen2009} consider three new indicators (firm size, prior return and stated motivation in the press release) and combine them with the book-to-market indicator in an undervaluation index (U-index).  Adding these additional indicators improves the predictability of excess returns, which is consistent with the undervaluation hypothesis.   In this paper we test whether two new indicators, standardized idiosyncratic volatility (measured by $1-R^2$) and volatility can improve our understanding of excess returns by combining them with the U-index of \cite{PeyerVermaelen2009}.

%An indicator that measures the competitive information advantage of managers is standardized idiosyncratic volatility, measured by $(1-R^2)$ where $R^2$ is, for example, that of a regression of the 6-months pre-announcement daily returns of the stock on the five factors of \cite{FamaFrench2015}.\footnote{Note that standardized idiosyncratic volatility differs from residual variance which is the measure of idiosyncratic volatility used in the volatility literature. Because of the high correlation between residual variance and total volatility (\Sexpr{round(100*cor(BUYBACK_DATA$DATASET$CRSP$pre_vol, BUYBACK_DATA$DATASET$CRSP$IVOL),2)}$\%$ in our sample) testing for the impact on volatility on stock returns is identical to testing for the impact of residual variance. Throughout this paper we use the term ''standardized idiosyncratic volatility'' to refer to $(1-R^2)$.} Indeed, it measures to what extent the volatility of stock returns is explained by company-specific (non-factor related) information [see also \citep{LiRajgopaletal2014}]. The hypothesis that managers are able to time the market is based on the assumption that managers have superior knowledge about company-specific information, not the overall market. Hence the prediction of the {\it information advantage hypothesis} is that buyback announcements of firms with high standardized idiosyncratic volatility $(1-R^2)$ will also generate higher long term excess returns.

%Adding (raw) volatility as a factor that explains excess returns can be justified by the \cite{Stambaugh2015} {\it costly arbitrage hypothesis} but also by the {\it option hypothesis} proposed by \cite{IkenberryVermaelen1996}:  a repurchase authorization is an option to take advantage of undervaluation and this option should be more valuable for high volatility stocks. In other words, the potential of managers to take advantage of undervaluation increases with the uncertainty about its fundamentals.  This option could be quite valuable as buyback authorizations are typically granted for several years and are often extended. If the market is efficient this option effect should be incorporated in the stock price at the time of the buyback announcement. Indeed, \cite{IkenberryVermaelen1996} find that short-term announcement returns are positively related to volatility. However, if markets underestimate the value of this option at the time of the buyback authorization, long term excess returns should be positively correlated with volatility. Note that, unlike the costly arbitrage hypothesis, this option hypothesis does not predict a relation between volatility and excess returns after equity issues which, unlike buyback authorizations, are firm commitments, not options.

%We find a significant positive relation between excess returns and volatility as well as standardized idiosyncratic volatility. Combining total volatility and standardized idiosyncratic volatility with the \cite{PeyerVermaelen2009} Undervaluation Index into an Enhanced Undervaluation Index (EU-Index) also improves the predictability of excess returns. In particular, during the four years following the buyback announcement, the high EU-Index portfolio generates an excess return of \Sexpr{round(EU_CALtable_bb["+48",13+6],2)}$\%$ per month with the Calendar Time event study method. Using the IRATS method the cumulative excess return reaches \Sexpr{round(EU_IRATStable_bb["+48",19],2)}$\%$ after 48 months.

%It remains a fact though that the buyback anomaly is to some extent a small firm anomaly, as also found by \cite{PeyerVermaelen2009}: indeed, we find that value-weighting all the events [as suggested by \cite{MitchellStafford2000}] makes the (Calendar Time event study method) alphas disappear. However, if we exclude the firms in the largest size quartile, the anomaly persists (also for the Calendar Time event study method).   This makes sense as small firms are more opaque and less followed by analysts. They also tend to be held more by insiders, so that in these case share buybacks are similar to insider buying with company funds. If anything, to increase the power of the test to detect mispricing and the possibility that managers (at least in some firms) can time the market based on their information advantage relative to investors, any weighting should be based on the inverse of size. The fact that excess returns disappear when we value weight provides further support for the hypothesis that excess returns are driven by information asymmetry for some firms. At the same time the negative relation between size and alpha may partly explain why this anomaly persists after 30 years and has attracted very little attention in the asset management industry.\footnote{For example, a Google search for ``buyback funds'' gives very few results: Powershare Buyback Achievers fund, KBC Buyback America, S\&P 500 Buyback ETF, Catalyst/Equity Compass Buyback Strategy fund, and PV Buyback USA. The first 3 funds focus on large caps after buyback completions although the academic research shows abnormal returns are more significant in small, under-priced, value stocks and the relevant event is not the completion of the buyback but the buyback authorization. We are also not aware of event-driven hedge funds that buy repurchasing firms and short equity issuers; typical event-driven strategies are for example based on M\&A arbitrage, capital structure arbitrage or on investing in distressed securities.}  Indeed because management fees are proportional to fund size one expects relatively less interest in anomalies concentrated in small caps or microcaps.

%Summarizing, this paper makes three fundamental contributions. First we show that the buyback anomaly, but not the equity issue one, survives after using the 5-factor model of Fama and French if buybacks are separated from equity issues and excess returns are calculated after the authorization date. Second, in agreement with \cite{Stambaugh2015}, we find support for the  hypothesis that firms announce buybacks because they are undervalued and because of costly arbitrage there will be a positive relation between volatility and returns. The results are also consistent with the \cite{IkenberryVermaelen1996} option hypothesis.  However, as the equity issue anomaly does not survive the Fama-French model we can't test the other prediction of \cite{Stambaugh2015}, i.e. that returns of overvalued firms will be negatively correlated with volatility and that this relation will be stronger for overvalued firms than undervalued firms.\footnote{Of course this does not mean that the \cite{Stambaugh2015} hypothesis is rejected as there could be other overvalued firms with a strong negative relation between excess returns and volatility.}  Finally, we show that adding volatility and standardized idiosyncratic volatility to the factors proposed by previous research as undervaluation proxies and combining them in an Enhanced Undervaluation index improves our forecasting ability of excess returns after buyback authorization announcements. As a theoretical argument can be made that both measures are proxies for the competitive information advantage of managers-insiders, this strengthens the support for the undervaluation and market timing hypothesis of some firms.

@@

For the buybacks we combined all open market repurchase announcements from both the SDC Repurchases data base and the SDC US mergers and acquisitions ($M\&A$) data base.
%\footnote{More information is available upon request. An interactive online tool to explore data variations and robustness analyses of all results in this paper, as well as all the source code, will also be available at \href{http://tevgeniou.github.io/BuybacksIssuers/}{tevgeniou.github.io/BuybacksIssuers.}}   
We ended up with a total of \Sexpr{cleanup_bb$Standardise$final_merge_data} repurchases events, out of which \Sexpr{cleanup_bb$Standardise$only_repurchase_deals} were only from the SDC Repurchases database, \Sexpr{cleanup_bb$Standardise$only_MA_deals} only from the SDC $M\&A$ database and \Sexpr{cleanup_bb$Standardise$overlap_deals} from both. Finally, we removed the following events: no CRSP returns or not all Compustat data available (\Sexpr{BUYBACK_DATA$cleanupNoPermno +  BUYBACK_DATA$cleanupNoCRSPdata + BUYBACK_DATA$cleanupMissingSomeValues} events); the percent of shares authorized was larger than \Sexpr{MAX_EVENT_SIZE}$\%$ (\Sexpr{cleanup_bb$biz_clean$EventSize_filter} events), or the closing price was less than \$\Sexpr{penny_stock_price_old} for events before 1995 or \$\Sexpr{penny_stock_price_recent} for the other (\Sexpr{cleanup_bb$biz_clean$Penny_stock_filter} events), or  the primary stock exchange was not the NYSE, the Nasdaq, or Amex (\Sexpr{cleanup_bb$biz_clean$major_markets_only} events).  Finally, we removed all events from firms in the Financial and Utilities sectors (\Sexpr{sum(BUYBACK_DATA$cleanupBIZ$Industry_filter)}  events).\footnote{We are using the industries from Kenneth French's Website. The Financial Sector consists of all firms with SIC code at the time of the buyback announcement that belonged in the ``Banks'' or ``Fin'' industries (SIC codes 6000 to 6300 and 6700 to 6799). The Utilities Sector consists of all firms with SIC code 4900 to 4942.}  At the end we are left with \Sexpr{length(cusip_bb)} buyback events made by \Sexpr{length(unique(cusip_bb))} firms. Table \ref{tbl:descriptive} summarizes the key data in this study. The average percent of shares authorized for these firms was \Sexpr{datasummaryBB["Percent authorized", "Mean"]}$\%$ (median of \Sexpr{datasummaryBB["Percent authorized", "Median"]}$\%$), the average Market Capitalization at announcement was \$\Sexpr{datasummaryBB["Market cap.", "Mean"]} Million (median of \$\Sexpr{datasummaryBB["Market cap.", "Median"]} Million), while the BE/ME was on average \Sexpr{datasummaryBB["BE/ME", "Mean"]} (median of \Sexpr{datasummaryBB["BE/ME", "Median"]}).

@@

Figure~\ref{fig:liveevents} shows the number of announcements per year in the sample period as well as the (standardized) level of the $S\&P$ 500. Buyback activity rises prior to stock market increases and tends to fall afterwards, especially during the financial crisis of 2008 when buyback announcements fell to a 15 year low. Note the structural decline in equity issues since 2000. A similar decline in IPOs is also observed by \cite{GaoRitterZhu2013}. 
%Figure \ref{fig:industryeventsBB} shows the number of buyback events per industry in our sample. The software, retail and semiconductor industries are the most active repurchasers.

@@

%\subsection{Share buybacks and Equity Issues in Isolation }
%\label{sec:AbnormalBuybacks}

@@

%The largest monthly excess return (\Sexpr{CAL_table_all_bb["+1","CAL 5F"]}$\%$) is earned by the portfolio that holds buyback stocks for one month  (not reported in Table \ref{tbl:priorBuybacks})  and the smallest excess return (\Sexpr{CAL_table_all_bb["+48","CAL 5F"]}$\%$) is earned by the portfolio that picks buybacks announced during the previous 48 months. This clearly shows that forming portfolios after buybacks are completed, as is done by measuring net issues in \cite{FamaFrench2016}, is introducing a downward bias as many repurchase programs are completed several months (sometimes years) after the buyback announcement. Waiting until the buyback is completed means missing the largest excess returns earned shortly after the buyback authorization. Finally, there are no statistically significant excess returns after equity issues, regardless whether we use the three or five-factor model.

%So far all our events are equally weighted. \cite{MitchellStafford2000} argue that events should be value weighted to test whether they represent an economically important anomaly. However, as pointed out by Loughran and Ritter (2000), small firms are more likely to be mispriced as they are less followed by analysts. t Hence managers of small firms are better able  to take advantage of mispricing than managers of  large firms. So value weighting the events would simply bias the results toward zero. And indeed, when we value-weight the events (see Table \ref{tbl:valuetable} Panel~A) long-term excess returns become statistically insignificantly different from zero when using the total sample of events. So the buyback anomaly is not economically important and does not challenge the basic premise that ``the market'' represented by a value-weighted index is priced correctly. Note, however, if we eliminate the firms with are in the top $25^{th}$ market capitalization percentile (see Table \ref{tbl:valuetable}, Panel~B) the results become again significant. These firms had market capitalization less than \$\Sexpr{round(max(BUYBACK_DATA$DATASET$CRSP$Market.Cap[BUYBACK_DATA$DATASET$SDC$Event.Date >= "2015-01-01" & BUYBACK_DATA$DATASET$CRSP$Market.Cap_score < 0.75]))} million in 2015.  The fact that the buyback anomaly is a small cap anomaly makes it more likely that excess returns are evidence that managers (at least in these firms) time the market based on their information advantage relative to investors.

@@

%Using IRATS, the 4 factor model produces significant positive excess returns of 12.66$\%$ after 4 years (t=5.24). With the calendar time method, the excess returns become less significant (t=1.81) but have the same magnitude (0.27$\%$ per month or 12.86$\%$ after 48 months). Clearly this is inconsistent with the idea that firms issue stock when they are overvalued.  Our results are largely consistent with \cite{FamaFrench2016}. Using the three-factor model, we find statistically significant long term (after 48 months) negative cumulative excess returns of -7.40% (t=-3.36). However, once we use the five-factor model as a benchmark, excess returns fall and become statistically insignificant after 48 months. Our results show that This indicates that when searching for anomalies, buybacks and equity issues should not be pooled in a “net issue” measure. Both events are clearly motivated by different reasons. As pointed out above, unlike buybacks, equity issues are firm commitments announced and completed at the same point in time, the buyer of the shares knows that the company is the issuer which makes trading against ”uninformed” investors more complicated and more controversial for a firm who wants to keep good relations with its long term investors. Note also that equity issues are typically preceded by large positive excess returns of around 37% in the 6 months prior to the equity issue. However, the lack of post announcement negative excess returns shows that this was not reflecting “irrational exuberance” but rather that, for example, these firms possibly experienced a substantial increase in growth opportunities and issued equity to finance them.

%shows a split according to volatility. We will return to this result next when we discuss the effect of volatility on returns in section 4.} using the same methodology as in Table~\ref{tbl:priorBuybacks}, Panel~A. Our results are largely consistent with \cite{FamaFrench2016}. Using the three-factor model, we find statistically significant long term (after 48 months) negative cumulative excess returns of \Sexpr{IRATS_table_all_iss["+48","CAR 3F"]}$\%$ (t=\Sexpr{IRATS_table_all_iss["+48",2]}). However, once we use the five-factor model as a benchmark, excess returns fall and become statistically insignificant after 48 months. This indicates that when searching for anomalies, buybacks and equity issues should not be pooled in a ``net issue'' measure. As pointed out above, unlike buybacks, equity issues are firm commitments announced and completed at the same point in time, the buyer of the shares knows that the company is the issuer which makes trading against ''uninformed'' investors more complicated and more controversial for a firm who wants to keep good relations with its long term investors. Note also that equity issues are typically preceded by large positive excess returns of around \Sexpr{floor(IRATS_table_all_iss[reported_times[1],4])}$\%$ in the 6 months prior to the equity issue. However, the lack of post announcement negative excess returns shows that this was not reflecting ``irrational exuberance'' but rather that, for example, these firms possibly experienced a substantial increase in growth opportunities and issued equity to finance them.

%Table~\ref{tbl:priorBuybacks}, Panel~A, shows the five-factor as well as the four-factor IRATS for high U-index and low U-index firms. The interesting conclusion is that using the five-factor model improves the predictive power of the U-index: high U-index firms earn 4 year excess returns of \Sexpr{IRATS_table_all_bb_undervaluation["+48","U:CAR5F"]}$\%$ (t=\Sexpr{IRATS_table_all_bb_undervaluation["+48",8]}) while low U-index firms only earn \Sexpr{IRATS_table_all_bb_undervaluation["+48","O:CAR5F"]}$\%$ (t=\Sexpr{IRATS_table_all_bb_undervaluation["+48",11]}), hence \Sexpr{IRATS_table_all_bb_undervaluation["+48","U:CAR5F"] - IRATS_table_all_bb_undervaluation["+48","O:CAR5F"]}$\%$ less than the high U-index ones. Starting from 24 months after the announcement, high U-index firms always beat low U-index firms. When we use the three-factor model, we find similar conclusions, but the results are weaker. For example after 48 months the high U-index firms now earn excess returns of \Sexpr{IRATS_table_all_bb_undervaluation["+48","U:CAR3F"]}$\%$, which is only \Sexpr{IRATS_table_all_bb_undervaluation["+48","U:CAR3F"] - IRATS_table_all_bb_undervaluation["+48","O:CAR3F"]}$\%$ higher than the low-U-index firms. Note that, consistent with \cite{PeyerVermaelen2009} the low U-index buyback stocks earn significant positive excess returns too. It is difficult to find a portfolio of buyback stocks that under-performs in the long run. So the term ``overvaluation'' should be interpreted with caution. The Undervaluation Index is a proxy for the likelihood that the buyback is driven by undervaluation. It does not imply that low U-index firms are overvalued. It means that for these firms the buyback is less likely to be driven by undervaluation, but by other reasons such as managing capital structure, avoiding dilution from executive stock options etc.

%Table~\ref{tbl:priorBuybacks}, Panel~B, shows that this conclusion holds when we use the Calendar Time method. High U-index stocks almost always beat low U-index stocks. As in the case of IRATS, the five-factor model improves the selectivity of the Undervaluation Index: low U-index now earn marginally significant excess returns after 48 months.

%Figure~\ref{fig:pastall} summarizes the IRATS results in more detail by showing the cumulative abnormal returns during the 6 month pre and 48 month post event period, using the five-factor model. Results are shown for the total sample of buybacks and equity issues, as well as for high and low U-index buyback samples.

@@

The results so far are based on a sample of all buyback and equity announcements over a thirty-year period. As the equity issue anomaly does not survive the \cite{FamaFrench2015} five-factor or the \cite{StambaughYuan2016} four-factor model, this section focuses on better understanding the buyback anomaly and uses the five-factor model as a benchmark.\footnote{All analyses below are also done for equity announcements. However, in agreement with the results in Section \ref{sec:PastFindingsBB}, we find no consistent/robust results for issuers. All issuers results are available upon request.}  We use the five factor model as Table \ref{tbl:priorBuybacks} shows that it better explains returns than the four-factor model. In other words, it is a more difficult benchmark to beat.\footnote{We also repeated the whole analysis below with the \cite{StambaughYuan2016} model but our main conclusions are essentially unchanged. Results are available upon request.} The purpose of this section is to test the robustness of this anomaly: has it become less important over time because markets have become more efficient? How sensitive is it to the length of the investment period? %Could the anomaly be industry-specific?

@@

%But considering the very high correlation between residual variance and total variance, their hypothesis would also predict a positive correlation between total volatility and future returns for undervalued stocks. Hence, perhaps total volatility is a better prediction of excess returns than (standardized) idiosyncratic volatility (defined as $1-R^2$) or the U-index of \cite{PeyerVermaelen2009}. Or perhaps volatility can be an additional, next to the U-index and standardized idiosyncratic volatility, indicator of the likelihood that the buyback is driven by undervaluation.\footnote{In contrast to the option hypothesis, \cite{Stambaugh2015} also predict a negative relation between volatility and returns for overvalued stocks. However, as with the Fama-French 5 factor model equity issuers don't appear to be overvalued, we can't test this ''second'' leg of their hypothesis.}

>>>>

if (generate_all_data == 1){
  source("create_project_data.R")
}
if (generate_rnw_data == 1){
  source("process_project_data.R")
} else {
  load("processed_project_data.Rdata")
}

@@

\doublespacing

@@

@@

@@

@@

@@

For the buybacks we combined all open market repurchase announcements from both the SDC Repurchases data base and the SDC US mergers and acquisitions ($M\&A$) data base. We ended up with a total of \Sexpr{cleanup_bb$Standardise$final_merge_data} repurchases events, out of which \Sexpr{cleanup_bb$Standardise$only_repurchase_deals} were only from the SDC Repurchases database, \Sexpr{cleanup_bb$Standardise$only_MA_deals} only from the SDC $M\&A$ database and \Sexpr{cleanup_bb$Standardise$overlap_deals} from both. Finally, we removed the following events: no CRSP returns or not all Compustat data available (\Sexpr{BUYBACK_DATA$cleanupNoPermno +  BUYBACK_DATA$cleanupNoCRSPdata + BUYBACK_DATA$cleanupMissingSomeValues} events); the percent of shares authorized was larger than \Sexpr{MAX_EVENT_SIZE}$\%$ (\Sexpr{cleanup_bb$biz_clean$EventSize_filter} events), or the closing price was less than \$\Sexpr{penny_stock_price_old} for events before 1995 or \$\Sexpr{penny_stock_price_recent} for the other (\Sexpr{cleanup_bb$biz_clean$Penny_stock_filter} events), or  the primary stock exchange was not the NYSE, the Nasdaq, or Amex (\Sexpr{cleanup_bb$biz_clean$major_markets_only} events).  Finally, we removed all events from firms in the Financial and Utilities sectors (\Sexpr{sum(BUYBACK_DATA$cleanupBIZ$Industry_filter)}  events).\footnote{We are using the industries from Kenneth French's Website. The Financial Sector consists of all firms with SIC code at the time of the buyback announcement that belonged in the ``Banks'' or ``Fin'' industries (SIC codes 6000 to 6300 and 6700 to 6799). The Utilities Sector consists of all firms with SIC code 4900 to 4942.}  At the end we are left with \Sexpr{length(cusip_bb)} buyback events made by \Sexpr{length(unique(cusip_bb))} firms. Table \ref{tbl:descriptive} summarizes the key data in this study. The average percent of shares authorized for these firms was \Sexpr{datasummaryBB["Percent authorized", "Mean"]}$\%$ (median of \Sexpr{datasummaryBB["Percent authorized", "Median"]}$\%$), the average Market Capitalization at announcement was \$\Sexpr{datasummaryBB["Market cap.", "Mean"]} Million (median of \$\Sexpr{datasummaryBB["Market cap.", "Median"]} Million), while the BE/ME was on average \Sexpr{datasummaryBB["BE/ME", "Mean"]} (median of \Sexpr{datasummaryBB["BE/ME", "Median"]}).

@@

Figure~\ref{fig:liveevents} shows the number of announcements per year in the sample period as well as the (standardized) level of the $S\&P$ 500. Buyback activity rises prior to stock market increases and tends to fall afterwards, especially during the financial crisis of 2008 when buyback announcements fell to a 15 year low. Note the structural decline in equity issues since 2000. A similar decline in IPOs is also observed by \cite{GaoRitterZhu2013}.

@@

@@

@@

@@

The results so far are based on a sample of all buyback and equity announcements over a thirty-year period. As the equity issue anomaly does not survive the \cite{FamaFrench2015} five-factor or the \cite{StambaughYuan2016} four-factor model, this section focuses on better understanding the buyback anomaly and uses the five-factor model as a benchmark.\footnote{All analyses below are also done for equity announcements. However, in agreement with the results in Section \ref{sec:PastFindingsBB}, we find no consistent/robust results for issuers. All issuers results are available upon request.}  We use the five factor model as Table \ref{tbl:priorBuybacks} shows that it better explains returns than the four-factor model. In other words, it is a more difficult benchmark to beat.\footnote{We also repeated the whole analysis below with the \cite{StambaughYuan2016} model but our main conclusions are essentially unchanged. Results are available upon request.} The purpose of this section is to test the robustness of this anomaly: has it become less important over time because markets have become more efficient? How sensitive is it to the length of the investment period?

@@
